Details of specific functions

### 2022-02-11 inst/gitbook/03_specific-tutorials.Rmd
Started the covar section wrtitng

What is a variance-covariance matrix
How to get them: MCMCglmm @MCMCglmm from @hadfield2010 (but see also @mulTree)

@@

@@

>>>>

Variance-covariance matrices are sometimes a useful way to summarise multidimensional data.
In fact, you can express the variation in your multidimensional dataset directly in terms of how your trait covary rather than simply the positions of your elements in the trait space.
Furthermore, variance-covariance matrices can be estimated from multidimensional in sometimes more useful ways that simply looking at the the data in your trait space.
This can be done by describing your data as hierarchical models like generalised linear mixed effect models (glmm).
For example, you might have a multidimensional dataset where your observations have a nested structure (e.g. they are part of the same phylogeny).
You can then analyse this data using a glmm with something like `my_data ~ observations + phylogeny + redisduals`.
For more info on these models [start here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model).
For more details on running these models, I suggest using the `MCMCglmm` package (@MCMCglmm) from @hadfield2010 (but see also @mulTree).

@@

Once you have a trait space and variance-covariance matrices output from the `MCMCglmm` model, you can use the function `MCMCglmm.subsets` to create a `"dispRity"` object that contains the classic `"dispRity"` data (the matrix, the subsets, etc...) but also a the new `$covar` element:

```{r}
## Loading the charadriiformes data
data(charadriiformes)
```

Here we using precaculated variance-covariance matrices from the charadriiformes dataset that contains a set of posteriors from a `MCMCglmm` model.
The model here was `data ~ traits + clade specific phylogenetic effect + global phylogenetic effect + residuals`.
We can retrieve the model information using the  [`MCMCglmm` utilities](#MCMCglmm-utilities) tools, namely the `MCMCglmm.levels` function to directly extract the terms names as used in the model and then build our `"dispRity"` object with the correct data, the posteriors and the correct term names:

```{r}
## The term names
model_terms <- MCMCglmm.levels(charadriiformes$posteriors)[1:4]
## Note that we're ignoring the 5th term of the model that's just the normal residuals

## The dispRity object
MCMCglmm.subsets(data = charadriiformes$data,
                 posteriors = charadriiformes$posteriors,
                 group = model_terms)
```

As you can see this creates a normal dispRity object with the information you are now familiar with.
However, we can be more fancy and provide more understandable names for the groups and provide the underlying phylogenetic structure used:

```{r}
## A fancier dispRity object
my_covar <- MCMCglmm.subsets(data = charadriiformes$data,
                             posteriors = charadriiformes$posteriors,
                             group = model_terms,
                             tree = charadriiformes$tree,
                             rename.groups = c(levels(charadriiformes$data$clade), "phylogeny"))
## Note that the group names is contained in the clade column of the charadriiformes dataset as factors
```

@@

One useful thing to do with these objects is then to visualise them in 2D.
Here we can use the `covar.plot` function (that has many different options that just `plot.dispRity` for plotting covar objects) to plot the trait space, the 95% confidence interval ellipses of the variance-covariance matrices and the major axes from these ellipses.
See the `?covar.plot` help page for all the options available:

```{r}
par(mfrow = c(2,2))
## The traitspace
covar.plot(my_covar, col = c("orange", "darkgreen", "blue"), main = "Trait space")
## The traitspace's variance-covariance mean ellipses
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Mean VCV ellipses",
           points = FALSE, ellipses = mean) 
## The traitspace's variance-covariance mean ellipses
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Mean major axes",
           points = FALSE, major.axes = mean)
## A bit of everything
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Ten random VCV matrices",
           points = TRUE, major.axes = TRUE, points.cex = 1/3, n = 10, ellipses = TRUE)
```
